Experience-Biased Symbolic Planning

### 2014-05-11 AAAI_AuthorKit/paper.tex
Grad AI version

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
% AAAI format packages
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
% Additional packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
% END Additional packages
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Experience-Biased Symbolic Planning)
/Author (Aram Ebtekar, Mike Phillips, Maxim Likhachev, Sven Koenig)}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Experience-Biased Symbolic Planning}
\author{Aram Ebtekar, Mike Phillips, Maxim Likhachev, Sven Koenig\\
Carnegie Mellon University\\
5000 Forbes Ave,\\
Pittsburgh, PA 15213\\
}
\maketitle
\begin{abstract}
\begin{quote}
Weighted A* search is the foundation for state-of-the-art domain-independent symbolic planners.
An important open challenge for these planners is to utilize experience with similar planning problems to speed up subsequent planning episodes.
Experience graphs have recently shown promise as a technique for plan reuse in the context of weighted A* searches in robotics, by biasing them towards a subgraph of the search space, typically chosen to consist of the edges used in previous plans.
In this paper, we demonstrate how to augment symbolic planners with
experience graphs, using the HSP STRIPS planner as a representative
example.
The resulting planners are able to trade off seamlessly
between plan quality and runtime and thus have advantages over past
plan-reuse techniques, such as planning by analogy.
We demonstrate
experimentally that our planner speeds up HSP by a factor of 2 or more on
many domains while achieving about the same plan quality.
\end{quote}
\end{abstract}

@@

In practice, even with a domain-independent means of generating heuristics, STRIPS problems can be very difficult.
Even humans have difficulty when faced with an unfamiliar kind of puzzle, though we get better with experience.
Thus, one might hope that a planning agent would similarly learn to generalize solutions from past planning experiences to related new instances.
A recent approach builds \textit{Experience graphs} to estimate the high-level connectivity of the free space in motion planning tasks.
The intuition is to remember previously generated paths so that, when a new start and goal is queried, a new path can be quickly generated by reusing subpaths from the E-graph.
However, E-graphs have never before been applied to high-level representations such as STRIPS. This is our present contribution.

@@

Many of today's state-of-the-art domain-independent STRIPS planners are based on the HSP planner, which we now describe.
It's common to compute heuristics by solving an easier, relaxed version of the original problem. In STRIPS, one might ignore the delete lists of operations.
Since having more atoms makes preconditions and the goal condition more likely to hold, this can only make the problem easier, so the resulting heuristic never overestimates the true cost.

@@

Let's say we want to compute a distance-to-$G$ heuristic from a state $S$. HSP estimates the cost to achieve an atom $a\in A$ from $S$ by $g_S(a) = $
\[\begin{cases} 0  &\mbox{if } a \in S
\\ \min_{op\mid a\in Add(op)} \left(g_S(Prec(op)) + Cost(op)\right)  &\mbox{if } a \notin S \end{cases}\]

The HSP heuristic is defined by
\[h^{HSP}(S) = g_S(G),\]
the estimated cost of achieving all atoms in $G$. Note that $g_S(\cdot)$ is evaluated on certain atom sets, namely $Prec(op)$ and $G$.
To keep the computations feasible, define $g_S(P) = \max_{p\in P} g_S(p)$ for atom sets $P$.
The resulting HSP-max heuristic is an underestimate; indeed, it satisfies a stronger property called \textbf{consistency}.

@@

\begin{thm}$h^E$ is $\epsilon^E$-consistent. [cite]\end{thm}

@@

This figure illustrates the idea. Dark edges have cost 3 and are in the E-graph. Light edges represent the jumps $h(S_i,S_j) = 2$ (with one exception: $h(S,G) = 5$), which we inflate by a factor $\epsilon^E=2$ to make 4 (or 10, respectively). Dijkstra's algorithm is used to compute the $h^E$ values shown alongside each node.

@@

Now we combine the ideas from HSP and E-graphs to construct a domain-independent STRIPS planner which learns from experience. In STRIPS, the consistent heuristic $h(S,S')$ is given by the HSP-max estimate $h^{HSP}_S(S')$. We can compute $h^{HSP}_S(S')$ for all $S,S'\in V^E$ in precisely $|V^E|$ runs of the dynamic programming we saw before: each run gives us $|V^E|$ of these results, yielding $(|V^E|)^2$ values in total.

@@

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\STATE Let $c(S,S')$ for $S,S'\in V^E$ represent E-graph edge costs.
\FORALL{$S \in V^E$}
\STATE ComputeG(S)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Run A* on the full graph from $I$ with heuristic $\epsilon h^E$:
\FOR{$S \notin V^E$ when generated}
\STATE ComputeG(S)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

ComputeG implements the dynamic programming computation of $g$-values. Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations. Multiplying everything together and ignoring cardinality signs for brevity, the runtime bound is $O(\mathbf{A^2O})$. Thus, each state generated costs $O(\mathbf{A^2O + V^E})$ time, and the preprocessing cost is asymptotically equivalent to preemptively generating every state in $V^E$. For comparison, standard HSP without E-graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate a state, and does not incur the cost of generating a state unless the A* encounters it.

@@

We augmented the HSP planner code [CITE] which won [AI COMPETITION IN 2000] with E-graphs. Our aim is to observe the effect of E-graphs on planning time given that we've already solved a closely related problem. To measure this, we run three searches for each problem instance. First, we run the original problem using standard HSP. Then, we generate a modified problem by moving the start and goal nodes randomly for X steps each. The second search solves the modified problem in the standard way. Finally, the third search solves the modified problem using as E-graph a random Y\% of the path returned from the first search. The reason for saving only a portion of the previous experience is to simulate the effects of disconnected E-graphs, as might arise after multiple searches. Our main comparisons will be between the number of nodes generated in the second and third searches. In practice, the times followed similar ratios to the number of nodes generated; as one would expect from the analyis in [PREVIOUS SECTION] since the E-graph resulting from Y\% of a single path is very small.

@@

@@

Generalization from experiences

Alternative E-graph edge-selection schemes (complete graph on $V^E$?)

@@

>>>>

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
% AAAI format packages
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
% Additional packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
% END Additional packages
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Experience-Biased Symbolic Planning)
/Author (Aram Ebtekar, Mike Phillips, Maxim Likhachev, Sven Koenig)
/Keywords (symbolic planning, E-graph, experience-biasedh, A* search, STRIPS, HSP)
}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Experience-Biased Symbolic Planning}
\author{Aram Ebtekar \and Mike Phillips \and Maxim Likhachev\\
Carnegie Mellon University\\
5000 Forbes Ave,\\
Pittsburgh, PA 15213\\
\And Sven Koenig\\
University of Southern California\\
Los Angeles, CA 90089
}
\maketitle
\begin{abstract}
\begin{quote}
Weighted A* search is the foundation for state-of-the-art domain-independent symbolic planners.
An important open challenge for these planners is to utilize experience with similar planning problems to speed up subsequent planning episodes.
Experience graphs have recently shown promise as a technique for plan reuse in the context of weighted A* searches in robotics, by biasing them towards a subgraph of the search space, typically chosen to consist of the edges used in previous plans.
In this paper, we demonstrate how to augment symbolic planners with
experience graphs, using the HSP STRIPS planner as a representative
example.
The resulting planners are able to trade off seamlessly
between plan quality and runtime and thus have advantages over past
plan-reuse techniques, such as planning by analogy.
We demonstrate
experimentally that our planner speeds up HSP by a factor of 2 or more on
many domains while achieving about the same plan quality.
\end{quote}
\end{abstract}

@@

In practice, even using state-of-the-art methods to generate heuristics on isolated planning instances, STRIPS problems can be very difficult to solve without domain knowledge.
Even humans have difficulty when faced with an unfamiliar kind of puzzle, though we get better with experience.
Thus, one might hope that a planning agent would similarly learn to generalize solutions from past planning experiences to related new instances.
A recent approach builds \textbf{Experience graphs} to estimate the high-level connectivity of the free space in motion planning tasks \cite{phillips2012graphs}.
The intuition is to remember previously generated paths so that, when a new start and goal is queried, a new path can be quickly generated by reusing subpaths from the E-graph.
However, E-graphs have never before been applied to high-level representations such as STRIPS. This is our present contribution.

@@

Many of today's state-of-the-art domain-independent STRIPS planners are based on the HSP planner \cite{bonet2001planning}, which we now describe.
It's common to compute heuristics by solving an easier, relaxed version of the original problem. In STRIPS, one might ignore the delete lists of operations.
Since having more atoms makes preconditions and the goal condition more likely to hold, this can only make the problem easier, so the resulting heuristic never overestimates the true cost.

@@

Let's say we want a heuristic estimate of the distance from $S$ to a goal that contains $G$. HSP estimates the cost to achieve an atom $a\in A$ from $S$ by $g_S(a) = $
\[\begin{cases} 0  &\mbox{if } a \in S
\\ \min_{op\mid a\in Add(op)} \left(g_S(Prec(op)) + Cost(op)\right)  &\mbox{if } a \notin S \end{cases}\]

The HSP heuristic is then defined by
\[h^{HSP}(S) = g_S(G),\]
the estimated cost of achieving all atoms in $G$. Note that $g_S(\cdot)$ is evaluated on certain atom sets, namely $Prec(op)$ and $G$.
To keep the computations feasible, define $g_S(P) = \max_{p\in P} g_S(p)$ for atom sets $P$.
The resulting HSP-max heuristic is an underestimate; indeed, it satisfies a stronger property called \textbf{consistency}.

@@

\begin{thm}$h^E$ is $\epsilon^E$-consistent. \cite{phillips2012graphs}\end{thm}

@@

The above figure illustrates the idea. Suppose the dark edges have cost 3 and make up the E-graph. Non-E-graph edges are not shown; instead, light edges represent the jump estimates $h(S_i,S_j) = 2$ (with one exception: $h(S,G) = 5$), which we inflate by a factor $\epsilon^E=2$ to make 4 (or 10, respectively). Dijkstra's algorithm follows the edges backward to compute the $h^E$ values shown alongside each node.

$G$ is on the E-graph, though this need not be the case in general. $V^E$ is the pentagon on the right, and its $h^E$ values are precomputed before the search. $h^E(S)$ is computed only when A* generates $S\notin V^E$.

@@

Now we combine the ideas from HSP and E-graphs to construct a domain-independent STRIPS planner which learns from experience. In STRIPS, a consistent heuristic $h(S,S')$ is given by the HSP-max estimate $h^{HSP}_S(S')$. We can compute $h^{HSP}_S(S')$ for all $S,S'\in V^E$ in precisely $|V^E|$ runs of the dynamic programming we saw before: each run gives us $|V^E|$ of these results, yielding $(|V^E|)^2$ values in total.

@@

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\STATE Let $c(S,S')$ for $S,S'\in V^E$ represent E-graph edge costs.
\FORALL{$S \in V^E$}
\STATE ComputeG(S)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Run A* on the full graph from $I$ with heuristic $\epsilon h^E$:
\FOR{$S \notin V^E$ when generated by A*}
\STATE ComputeG(S)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

ComputeG implements the dynamic programming computation of $g$-values. Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations. Multiplying everything together and ignoring cardinality signs for brevity, the runtime bound is $O(\mathbf{A^2O})$. Thus, each state generated costs $O(\mathbf{A^2O + V^E})$ time. The preprocessing cost is $O(\mathbf{A^2OV^E + (V^E)^2})$, asymptotically equivalent to preemptively generating every state in $V^E$. For comparison, standard HSP without E-graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate a state, and does not incur the cost of generating a state unless the A* encounters it.

@@

We augmented the HSP planner code \cite{bonet2001planning} which won [AI COMPETITION IN 2000] with E-graphs. Our aim is to observe the effect of E-graphs on planning time given that we've already solved a closely related problem. To measure this, we run three searches for each problem instance. First, we run the original problem using standard HSP. Then, we generate a modified problem by moving the start and goal nodes randomly for X steps each. The second search solves the modified problem in the standard way. Finally, the third search solves the modified problem using as E-graph a random Y\% of the path obtained from the first search. The reason for saving only a portion of the previous experience is to simulate the effects of disconnected E-graphs, as might arise after multiple searches. Our main comparisons will be between the number of nodes generated in the second and third searches. In practice, the times followed similar ratios to the number of nodes generated; as one would expect from the analyis in [PREVIOUS SECTION] since the E-graph resulting from Y\% of a single path is very small.

@@

\includegraphics[scale=0.4]{AIPlot.png}

The plot shows speedup (i.e. ratio of nodes generated in the second and third searches) against a measure of problem size (i.e. nodes generated in the second search). Planning times followed similar trends as node generation counts, as one would expect from the analyis since the E-graph resulting from a half-path is very small. We see considerable speedups in many domains, while the plan costs stayed roughly the same (in fact, they became slightly shorter on average). Many domains saw no discernable effect, but none were significantly hurt by E-graphs.

@@

We saw that by reusing pieces from past plans, planning time can often be reduced. Many interesting directions remain for future investigation. The E-graph cannot be allowed to grow without bound over countless planning episodes, so a better scheme is needed for expansion and pruning as the agent learns.

Distinct states in STRIPS can actually be very similar, differing only by a few irrelevant atoms; hence, experiences should be transferable to inexact matches.

@@

\bibliographystyle{aaai}
\bibliography{paper.bib}
